Collaborative Multi-output Gaussian Processes

### 2014-06-11 paper/experiments.tex
perhaps final edits
move table for better format
1. add author names 2. fix references 3. some typos

\begin{figure*}
\centering
\begin{tabular}{cccc}
\includegraphics[scale=0.2]{figures/toy-slfm-y1.eps} &
\includegraphics[scale=0.2]{figures/toy-svigp-y1.eps} &
\includegraphics[scale=0.2]{figures/toy-slfm-y2.eps} &
\includegraphics[scale=0.2]{figures/toy-svigp-y2.eps}
\end{tabular}
\caption{Real data and predictive distributions of by COGP (first and third figure) and independent GPs using stochastic variational inference (second and last figure) for the toy problem. Solid black line: predictive mean; grey bar: two standard deviations; magenta dots: real observations; blue dots: missing data. The black crosses show the locations of the inducing inputs. By sharing inducing points across the outputs, COGP accurately interpolates the missing function values.}
\label{fig:toy}
\end{figure*}

\subsection{FOREIGN EXCHANGE RATE PREDICTION}
The first real world application we consider is to predict the foreign exchange rate w.r.t the US dollar of the top 10 international currencies (CAD, EUR, JPY, GBP, CHF, AUD, HKD, NZD, KRW, and MXN) and 3 precious metals (gold, silver, and platinum)\footnote{Data is available at http://fx.sauder.ubc.ca}. 
The setting of our experiment described here is identical to that in \citet{alvarez2010efficient}.
The dataset consists of all the data available for the 251 working days in the year of 2007.
There are 9, 8, and 42 days of missing values  for gold, silver, and platinum, respectively.
We remove from the data the exchange rate of CAD on days 50--100, JPY on day 100--150, and AUD on day 150--200.
Note that these 3 currencies are from very different geographical locations, making the problem more interesting. 
The 153 points are used for testing, and the remaining 3051 data points are used for training.
Since the missing data corresponds to long contiguous sections, the objective here is to evaluate the capacity of the model to impute the missing currency values based on other currencies.
%todo: batch size, learn rate (maybe at the beginning of experiments)

>>>>

\begin{figure*}
\centering
\begin{tabular}{cccc}
\includegraphics[scale=0.2]{figures/toy-slfm-y1.eps} &
\includegraphics[scale=0.2]{figures/toy-svigp-y1.eps} &
\includegraphics[scale=0.2]{figures/toy-slfm-y2.eps} &
\includegraphics[scale=0.2]{figures/toy-svigp-y2.eps}
\end{tabular}
\caption{Simulated data and predictive distributions of by COGP (first and third figure) and independent GPs using stochastic variational inference (second and last figure) for the toy problem. Solid black line: predictive mean; grey bar: two standard deviations; magenta dots: real observations; blue dots: missing data. The black crosses show the locations of the inducing inputs. By sharing inducing points across the outputs, COGP accurately interpolates the missing function values.}
\label{fig:toy}
\end{figure*}

\subsection{FOREIGN EXCHANGE RATE PREDICTION}
The first real world application we consider is to predict the foreign exchange rate w.r.t the US dollar of the top 10 international currencies (CAD, EUR, JPY, GBP, CHF, AUD, HKD, NZD, KRW, and MXN) and 3 precious metals (gold, silver, and platinum)\footnote{Data is available at http://fx.sauder.ubc.ca}. 
The setting of our experiment described here is identical to that in \citep{alvarez2010efficient}.
The dataset consists of all the data available for the 251 working days in the year of 2007.
There are 9, 8, and 42 days of missing values  for gold, silver, and platinum, respectively.
We remove from the data the exchange rate of CAD on days 50--100, JPY on day 100--150, and AUD on day 150--200.
Note that these 3 currencies are from very different geographical locations, making the problem more interesting. 
The 153 points are used for testing, and the remaining 3051 data points are used for training.
Since the missing data corresponds to long contiguous sections, the objective here is to evaluate the capacity of the model to impute the missing currency values based on other currencies.
%todo: batch size, learn rate (maybe at the beginning of experiments)

### 2014-06-11 paper/inference.tex
1. add author names 2. fix references 3. some typos

\subsection{VARIATIONAL LOWER BOUND \label{sec:variationalLowerBound}}
\newcommand{\ug}{\u_g}
\newcommand{\uh}{\u^h}
\newcommand{\mgj}{\m_j}
\newcommand{\mhi}{\m^h_i}
\newcommand{\Sgj}{\S_j}
\newcommand{\Shi}{\S^h_i}
Our inference task is to find the posterior $p(\g, \h, \u, \v | \y)$, for which we use variational inference \citep{jordan-variational-99}. 
We first observe that true the posterior distribution can be written as:
\begin{align}
p(\g, \h, \u, \v | \y) = p(\g | \u, \y) p(\h | \v, \y) p(\u, \v | \y).
\end{align}
Here we recall the modeling assumption that each set of the inducing variables is the \emph{sufficient statistics} of the corresponding latent process. 
This motivates replacing the true posteriors over $\g$ and $\h$ with their conditional distributions given the inducing variables, leading to a distribution that of the form:
\begin{align}
q(\g, \h, \u, \v | \y)
= p(\g|\u) p(\h|\v) q(\u,\v),
\end{align}
with
\begin{align}
q(\u,\v) &= \prod_{j=1}^Q \underbrace{\Normal(\u_j; \mgj, \Sgj)}_{q(\u_j)} \prod_{i=1}^P  \underbrace{\Normal(\v_i; \mhi, \Shi)}_{q(\v_i)}.
\end{align}
This technique has been used by \citet{titsias2009variational} and \citet{hensmangaussian} 
to derive variational inference algorithms for the single output case.
Since the conditionals $p(\g | \u)$ and $p(\h | \v)$ are known (Equations \eqref{eq:gu} and \eqref{eq:hv}), 
we only need to 
learn $q(\u, \v)$ so as to minimize the divergence between the approximate posterior and
the true posteriors.
The quality of aproximation depends entirely on the posterior over the inducing variables, thus underlining their pivotal role in the model as previously discussed.

@@

The novelty of the  variational lower bound in Equation  \eqref{eq:L} is that it decomposes across both inputs and outputs.
This enables the use of stochastic optimization methods, which 
allow the model to handle  ver large datasets for which existing GP-based multi-output models are simply 
impractical.  
%
\subsection{STOCHASTIC VARIATIONAL INFERENCE \label{sec:svi}}
\newcommand{\oi}{\vec{o}_i}
% some notation

>>>>

\subsection{VARIATIONAL LOWER BOUND \label{sec:variationalLowerBound}}
\newcommand{\ug}{\u_g}
\newcommand{\uh}{\u^h}
\newcommand{\mgj}{\m_j}
\newcommand{\mhi}{\m^h_i}
\newcommand{\Sgj}{\S_j}
\newcommand{\Shi}{\S^h_i}
Our inference task is to find the posterior $p(\g, \h, \u, \v | \y)$, for which we use variational inference \citep{jordan-variational-99}. 
We first observe that the true posterior distribution can be written as:
\begin{align}
p(\g, \h, \u, \v | \y) = p(\g | \u, \y) p(\h | \v, \y) p(\u, \v | \y).
\end{align}
Here we recall the modeling assumption that each set of the inducing variables is the \emph{sufficient statistics} of the corresponding latent process. 
This motivates replacing the true posteriors over $\g$ and $\h$ with their conditional distributions given the inducing variables, leading to a distribution of the form:
\begin{align}
q(\g, \h, \u, \v | \y)
= p(\g|\u) p(\h|\v) q(\u,\v),
\end{align}
with
\begin{align}
q(\u,\v) &= \prod_{j=1}^Q \underbrace{\Normal(\u_j; \mgj, \Sgj)}_{q(\u_j)} \prod_{i=1}^P  \underbrace{\Normal(\v_i; \mhi, \Shi)}_{q(\v_i)}.
\end{align}
This technique has been used by \citet{titsias2009variational} and \citet{hensmangaussian} 
to derive variational inference algorithms for the single output case.
Since the conditionals $p(\g | \u)$ and $p(\h | \v)$ are known (Equations \eqref{eq:gu} and \eqref{eq:hv}), 
we only need to 
learn $q(\u, \v)$ so as to minimize the divergence between the approximate posterior and
the true posteriors.
The quality of approximation depends entirely on the posterior over the inducing variables, thus underlining their pivotal role in the model as previously discussed.

@@

The novelty of the  variational lower bound in Equation  \eqref{eq:L} is that it decomposes across both inputs and outputs.
This enables the use of stochastic optimization methods, which 
allow the model to handle  very large datasets for which existing GP-based multi-output models are simply 
impractical.  
%
\subsection{STOCHASTIC VARIATIONAL INFERENCE \label{sec:svi}}
\newcommand{\oi}{\vec{o}_i}
% some notation
